```{r 10_setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, eval=FALSE, fig.align="center")
```

# (PART) Estimation {-}

# Applied Analysis: Regression

## Learning Goals {-}

- Conduct and interpret results from a regression analysis to estimate average causal effects and subgroup average causal effects

<br>

Slides from today are available [here](https://docs.google.com/presentation/d/1kDpUsGSmvGP29rw2MB5VCErWz1oR0ez9YohzfEqZUwI/edit?usp=sharing).


<br><br><br>


## Analysis {-}

**You can download a template RMarkdown file to start from [here](template_rmds/10-applied-regression.Rmd).**


### Data context and research questions {-}

```{r}
library(tidyverse)
library(splines) # Make sure that the splines package is installed before running

tenure <- read_csv("aer_primarysample.csv")

tenure <- tenure %>%
    mutate(
        faculty = ifelse(faculty_miss==1, NA, faculty),
        revenue = ifelse(revenue_miss==1, NA, revenue),
        female_ratio = ifelse(female_ratio_miss==1, NA, female_ratio),
        full_ratio = ifelse(full_ratio_miss==1, NA, full_ratio),
        phd_rank = ifelse(phd_rank_miss==1, NA, phd_rank),
        max_csstops = ifelse(max_csstops_miss==1, NA, max_csstops)
    )
```

**Research questions:** Does a policy that delays the tenure process (for new parents) actually help tenure outcomes? Does it help men and women equally?

The data and full codebook are available [on Moodle](https://moodle.macalester.edu/mod/folder/view.php?id=36846) (`aer_primarysample.csv` and `ReadMe.pdf`). An abbreviated codebook is below:

**Treatment and primary outcome:**

- `gncs`: Is a gender-neutral clock-stopping policy in place at this person's school? (treatment)
- `tenure_policy_school`: Did this person get tenure at the university with the clock-stopping policy? (primary outcome)

<br>

**Additional outcomes (mediators):**

- `top_pubsX`: cumulative number of peer-reviewed publications in top-5 journals by year X (3, 5, 7, and 9) since PhD completion 
- `PUBSX`: cumulative number of non-top-5 peer-reviewed publications by year X (3, 5 7, and 9) since PhD completion

<br>

**Additional covariates:**

- `female`: Is the faculty member a female?
- `pol_u`: policy university identifier (values are randomized for privacy reasons)
- `pol_job_start`: identifier for year the job at the policy university started (values are randomized for privacy reasons)
- `phd_rank`: 1-5 categorical variable of PhD program tier based on placements into the top-50 departments in our sample
- `phd_rank_miss`: indicator for missing PhD information
- `post_doc`: indicator for doing a post-doc before the first tenure-track position
- `ug_students`: number of undergraduate students at the university, in thousands 
- `grad_students`: number of graduate students at the university, in thousands 
- `faculty`: number of faculty at the university (all disciplines), in hundreds 
- `full_av_salary`: average salary of full professors at the university, in thousands 
- `assist_av_salary`: average salary of assistant professors at the university, in thousands, 
- `revenue`: annual revenue of the university, in 10,000,000s 
- `female_ratio`: fraction of the faculty at the university who are female 
- `full_ratio`: fraction of the faculty at the university who are full professors
- `RANK`: equal to 1 for top-10 departments and 2 for all other departments
- `max_csstops`: number of children born within five years of PhD completion



### Part 1: Finalize adjustment set {-}

If you constructed your causal graph on DAGitty last class, paste your model code into [this Google Doc](https://docs.google.com/document/d/16NMaOKVHDaFy43tuHYjIaYRqSM3ct0Op97kC8uxT9es/edit?usp=sharing), and report the variables needed to achieve conditional exchangeability.

- Pick one graph to focus on for the remainder of this analysis. Open [DAGitty](http://www.dagitty.net/dags.html) and paste the model code from the Google Doc into the "Model code" pane on the right. Click "Update DAG" to view the graph.

- Spend a *short* amount of time reviewing this graph to see if you generally agree with the relationships depicted. Make any updates that you think are sensible. (Again, keep this short - in practice, this should take weeks of thorough literature review and expert consultation.)

- If any variables do not directly correspond to available variables, find proxies for them. (Several variables may end up serving as a proxy for one unmeasured variable on your graph.) Based on this, decide on your final adjustment set.



<br><br>



### Part 2: Exploratory analysis {-}

Construct visualizations to inform your regression model specification. Your visualizations should:

- Explore nonlinear relationships between conditioning variables and the outcome
- Explore an interaction between **one pair** of conditioning variables. (To keep our analysis short)

**Note:** All of the variables are coded as numeric, so if you need to make bar plots, you may need to wrap the variable name inside `factor()`, e.g., `factor(female)`.

The code below will be useful for exploring nonlinearity in covariate-outcome relationships. The blue smooth reflects observed data trends, and the red smooth shows the predictions from a logistic regression model with a specific model formula.

- Formula `y ~ x`: Covariate is included as a linear term
- Formula `y ~ ns(x,3)`: Covariate is modeled with a flexible nonlinear function (a "spline" with 3 degrees of freedom--don't worry about the details of this)

The model predictions should generally align with observed data trends. If they don't line up well using formula `y ~ x`, update the formula to `y ~ ns(x, 3)` to see if there is an improvement.

```{r}
ggplot(data, aes(x = covariate, y = tenure_policy_school)) +
    geom_point() +
    geom_smooth(se = FALSE, color = "blue", method = "loess") +
    geom_smooth(formula = INSERT_MODEL_FORMULA, method = "glm",
        method.args = list(family="binomial"),
        se = FALSE, color = "red"
    )
```


<br><br>


### Part 3: Modeling {-}

Fit a model to estimate the overall average causal effect and another to estimate the subgroup average causal effects for males and females. (Recall: `A*B` in a model formula creates an interaction between `A` and `B`.)

For both models, display the summarized output table and 95% confidence intervals for the coefficients of interest.

Interpret the coefficients of interest on the natural (exponentiated) scale.

```{r}
mod_overall <- glm(INSERT_MODEL_FORMULA, data = tenure, family = "binomial")
mod_bygender <- glm(INSERT_MODEL_FORMULA, data = tenure, family = "binomial")

summary(mod_overall)
summary(mod_bygender)

# On the log scale
confint(mod_overall)
confint(mod_bygender)

# On the natural scale
confint(mod_overall) %>% exp()
confint(mod_bygender) %>% exp()
```


<br><br>


### Part 4: Interpretation and discussion {-}

- Using both the confidence intervals and effect magnitudes, discuss the results of your analysis in a contextually meaningful way. (Tie these results back to the research questions.)

- Discuss limitations of your analysis by acknowledging the points in the analysis process that were most uncertain.


<br><br>


## Simulation study planning {-}

Describe in detail how you would implement a simulation study to study the impact of model misspecfication (incorrect form) on the accuracy of regression for estimating causal effects. It may help to draw a specific small causal graph as the basis of your simulation.
