---
title: "Applied Analysis: Inverse Probability Weighting"
output: 
  html_document:
    toc: true
    toc_float: true
---

## Data context and research questions

**Research questions:** Does a policy that delays the tenure process (for new parents) actually help tenure outcomes? Does it help men and women equally?

The data and full codebook are available [on Moodle](https://moodle.macalester.edu/mod/folder/view.php?id=36846) (`aer_primarysample.csv` and `ReadMe.pdf`). An abbreviated codebook is below:

**Treatment and primary outcome:**

- `gncs`: Is a gender-neutral clock-stopping policy in place at this person's school? (treatment)
- `tenure_policy_school`: Did this person get tenure at the university with the clock-stopping policy? (primary outcome)

**Additional outcomes (mediators):**

- `top_pubsX`: cumulative number of peer-reviewed publications in top-5 journals by year X (3, 5, 7, and 9) since PhD completion 
- `PUBSX`: cumulative number of non-top-5 peer-reviewed publications by year X (3, 5 7, and 9) since PhD completion

**Additional covariates:**

- `female`: Is the faculty member a female?
- `pol_u`: policy university identifier (values are randomized for privacy reasons)
- `pol_job_start`: identifier for year the job at the policy university started (values are randomized for privacy reasons)
- `phd_rank`: 1-5 categorical variable of PhD program tier based on placements into the top-50 departments in our sample
- `phd_rank_miss`: indicator for missing PhD information
- `post_doc`: indicator for doing a post-doc before the first tenure-track position
- `ug_students`: number of undergraduate students at the university, in thousands 
- `grad_students`: number of graduate students at the university, in thousands 
- `faculty`: number of faculty at the university (all disciplines), in hundreds 
- `full_av_salary`: average salary of full professors at the university, in thousands 
- `assist_av_salary`: average salary of assistant professors at the university, in thousands, 
- `revenue`: annual revenue of the university, in 10,000,000s 
- `female_ratio`: fraction of the faculty at the university who are female 
- `full_ratio`: fraction of the faculty at the university who are full professors
- `RANK`: equal to 1 for top-10 departments and 2 for all other departments
- `max_csstops`: number of children born within five years of PhD completion





### Loading data and packages

You'll need to install the `survey` package first.

```{r}
library(survey) # install.packages("survey")
library(tidyverse)
library(splines) # Make sure that the splines package is installed before running

tenure <- read_csv("aer_primarysample.csv")

tenure <- tenure %>%
    mutate(
        faculty = ifelse(faculty_miss==1, NA, faculty),
        revenue = ifelse(revenue_miss==1, NA, revenue),
        female_ratio = ifelse(female_ratio_miss==1, NA, female_ratio),
        full_ratio = ifelse(full_ratio_miss==1, NA, full_ratio),
        phd_rank = ifelse(phd_rank_miss==1, NA, phd_rank),
        max_csstops = ifelse(max_csstops_miss==1, NA, max_csstops)
    )
```





## Part 1: Review outcome regression results

Previously we used outcome regression to estimate the overall effect of tenure clock stopping policies and the effects for males and females. Models from one of our analyses are displayed here.

Review the results of this analysis by interpreting the relevant coefficients and confidence intervals on the odds ratio scale.

```{r}
# Overall ACE
mod_overall <- glm(tenure_policy_school ~ gncs + ns(female_ratio, 3) + ns(full_ratio, 3) + factor(max_csstops)*female, data = tenure, family = "binomial")
summary(mod_overall)
confint(mod_overall)

# Subgroup ACEs
mod_subgroup <- glm(tenure_policy_school ~ gncs*female + ns(female_ratio, 3) + ns(full_ratio, 3) + factor(max_csstops)*female, data = tenure, family = "binomial")
summary(mod_subgroup)
confint(mod_subgroup)
```





## Part 2: Propensity score modeling

We will assume that the `revenue`, `female_ratio`, `full_ratio`, and `max_csstops` are sufficient (proxies for) variables needed to achieve conditional exchangeability. (`female` was included in the outcome regresion models for assessing subgroup effects but wasn't determined to be part of a noncausal path in our causal graph analysis.)

- Fit an appropriate propensity score model called `ps_mod`. Use visualizations to inform the construction of your model. (Just focus on appropriately handling nonlinearity.)

- Use your model to compute appropriate weights, and add these weights to the dataset.

- Use "before and after" visualizations to compare balance of key variables before and after weighting. You'll need to use `factor(gncs)` to make your visualizations--this forces R to view `gncs` as a categorical variable.
    - Note: If any distributions look dissimilar after weighting, this is an indication that the propensity score model may be misspecified. We won't address this in our analysis today for time reasons, but in practice, more flexible modeling techniques would be explored.

Coding notes:

The code below will be useful for exploring nonlinearity in covariate-treatment relationships. The blue smooth reflects observed data trends, and the red smooth shows the predictions from a logistic regression model with a specific model formula.

- Formula `y ~ x`: Covariate is included as a linear term
- Formula `y ~ ns(x,3)`: Covariate is modeled with a flexible nonlinear function (a "spline" with 3 degrees of freedom--don't worry about the details of this)

The model predictions should generally align with observed data trends. If they don't line up well using formula `y ~ x`, update the formula to `y ~ ns(x, 3)` to see if there is an improvement.

```{r}
ggplot(DATA, aes(x = COVARIATE, y = OUTCOME)) +
    geom_point() +
    geom_smooth(se = FALSE, color = "blue", method = "loess") +
    geom_smooth(formula = INSERT_MODEL_FORMULA, method = "glm",
        method.args = list(family="binomial"),
        se = FALSE, color = "red"
    )
```



The (incomplete) code below is useful for computing appropriate weights:

```{r}
DATA <- DATA %>%
    mutate(
        ps = predict(ps_mod, newdata = DATA, type = "response"),
        ipw = case_when(
            TREAT_VAR==1 ~ ???,
            TREAT_VAR==0 ~ ???
        )
    )
```



You can incorporate weights into most `ggplot2` figures by adding `weight` to the `aes`thetics:

```{r}
aes(..., weight = ip_weight)
```





## Part 3: Modeling with IP weights

We can use the IP weights constructed above in a weighted regression model to estimate causal effects.

- Note: The `svydesign()` and `svyglm()` functions from the `survey` package are used to ensure that the weights are treated appropriately. (An updated ("robust") standard error calculation method is used to appropriately account for the fact that the IP weights are estimated and have uncertainty.)

- Note that we are removing cases with missing values for the weights. (These cases had missing values in the covariates used in the propensity score model.) You'll think about whether this biases the analysis in Part 4.

- The code for estimating the overall ACE is complete. Adapt this code to estimate subgroup ACEs in males and females.

- Using both the confidence intervals and effect magnitudes, discuss the results of your analysis in a contextually meaningful way. (Tie these results back to the research questions.)

- How do confidence interval widths compare to those from outcome regression?

```{r}
# Remove cases with missing weights
tenure_subs <- tenure %>%
    filter(!is.na(ipw))

# Set up information about weights
design <- svydesign(ids = ~0, weights = tenure_subs$ipw, data = tenure_subs)

# Fit a marginal structural model to estimate overall ACE
overall_fit <- svyglm(
    tenure_policy_school ~ gncs,
    data = tenure_subs,
    design = design,
    family = "quasibinomial"
)
summary(overall_fit)
confint(overall_fit)

# Adapt the svyglm() code above to include an interaction between treatment and female
subgroup_fit <- 
```





## Part 4: Causal graphs and missing data

Challenge! How might we use causal graphs to analyze the impact of dropping missing data (a complete case analysis)? Can we use ideas from representing selection bias with causal graphs?


