---
title: "IPW: Simulation"
output: 
  html_document:
    toc: true
    toc_float: true
---

# Exercise 1: Simulate

We'll simulate data from the causal graph below where $Z$, $W$, and $A$ are binary, and $Y$ is quantitative. Use the structural equations below for $A$ and $Y$, and store your simulated data as `sim_data`.

- $\log(\mathrm{odds}(A = 1)) = -1 + 0.4 Z + 0.4 W + 0.9 Z*W$
- $Y$ follows a normal distribution with mean $10 + 5A + 6Z + 7W$ and standard deviation 2.

Based on the structural equation for $Y$, what is the true ACE? Based on the causal graph, what variables are needed to achieve conditional exchangeability?

```{r 11_ex1, echo=FALSE, eval=TRUE, fig.width=12, fig.height=4, fig.align="center"}
library(dagitty)
dag <- dagitty("
dag {
bb=\"0,0,1,1\"
A [exposure,pos=\"0.300,0.500\"]
W [pos=\"0.550,0.280\"]
Y [outcome,pos=\"0.600,0.500\"]
Z [pos=\"0.350,0.280\"]
A -> Y
W -> A
W -> Y
Z -> A
Z -> Y
}
")
plot(dag)
```

```{r}
set.seed(451)
n <- 10000
Z <- rbinom(n, size = 1, prob = 0.5)
W <- rbinom(n, size = 1, prob = 0.5)

log_odds_A <- 
odds_A <- 
p_A <- 
A <- 

noise_Y <- rnorm(___)
mean_Y <- 
Y <- mean_Y + noise_Y

sim_data <- data.frame(Z = factor(Z), W = factor(W), A = factor(A), Y)
```





## Exercise 2: Inverse probability weighting

Here, we'll check the performance of inverse probability weighting for estimating the average causal effect.

### Part a

The accuracy of IPW depends on having the right model for treatment as a function of the variables needed to achieve conditional exchangeability.

- Based on our simulation, what is the correct logistic regression model that should be fit? Fit this model as `ps_mod_complex`.
- Also fit a wrong model called `ps_mod_simple` that uses the formula `A ~ Z + W`. (In this way, we can explore the impact of model misspecification.)

```{r}
ps_mod_simple <- glm(FORMULA, data = sim_data, family = "binomial")
ps_mod_complex <- glm(FORMULA, data = sim_data, family = "binomial")
```

### Part b

The code below uses your models to compute (right and wrong) IP weights.

- `predict(logistic_mod, newdata = data_to_make_predictions_for, type = "response")` is used to predict probabilities from a logistic regression model. (Without `type = "response"`, log-odds are computed.)
- Add comments to this code to document these different pieces, and check in with the instructor if you have questions.

```{r}
sim_data <- sim_data %>%
    mutate(
        ps_simple = predict(ps_mod_simple, newdata = sim_data, type = "response"),
        ipw_simple = case_when(
            A==1 ~ 1/ps_simple,
            A==0 ~ 1/(1-ps_simple)
        )
    ) %>%
    mutate(
        ps_complex = predict(ps_mod_complex, newdata = sim_data, type = "response"),
        ipw_complex = case_when(
            A==1 ~ 1/ps_complex,
            A==0 ~ 1/(1-ps_complex)
        )
    )
```

### Part c

Fit an ordinary regression model `Y ~ A` that ignores the IP weights. Is the estimated ACE what you expected?

### Part d

Incorporate the IP weights into your analysis by modifying your model from part c as below. (As discussed in 12.4 of WHATIF, this weighting fits a **marginal structural model (MSM)** that allows us to directly estimate the ACE.) Is the estimated ACE what you expected?

```{r}
lm(..., data = ..., weights = ipw_simple)
lm(..., data = ..., weights = ipw_complex)
```


**Note:** Using `weights = ...` in `glm(..., family = "binomial")` does not work exactly the way we want it to. (It works as we want for `lm()`.) Going forward, we will use a specialized package (the `survey` package) for dealing with weights.





## Exercise 3: Balance checking

### Part a

What is the key differentiator between regression and inverse probability weighting?

### Part b

Let's verify that unique property of IP weighting. Make plots to show the relationship between $Z$ and $A$ and between $W$ and $A$ in the original, unweighted data.

### Part c

Modify your plot to incorporate the IP weights by adding the following to `aes()`. What property of IP weighting does this show?

```{r}
aes(..., weight = ipw_complex)
```
